---
title: "linear regression"
author: "Arash Hatamirad, ahatamirad@gmail.com"
date: "9/1/2021"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(root.dir = "C:/Users/Sanaz/Documents/All_MSDA_Projects and Sources/R Ptojects/Algorithm I/HW03/AlgorithmI-HW3")

# Install required packages

# install.packages("MASS")
# install.packages("fBasics")
# install.packages("moments")
# install.packages("dplyr")
# install.packages("car")
#install.packages("DescTools")


library(MASS)
library(fBasics)
library(moments)
library(dplyr)
library(car)
library(DescTools)
library(olsrr)


#Data: heartbpchol.csv for Exercise1, 
# bupa.csv for Exercise 2, 
# psych.csv for Exercise3, 
# cars_new.csv for Exercise 4
```


## linear regression  


We will use heart.csv dataset. Below is brief summary of variables in heart.csv.  

- Weight: subject’s weight  
- Systolic: top number in a blood pressure reading, indicating the blood pressure level when the heart contracts  
- Diastolic: bottom number in a blood pressure reading, indicating the blood pressure level when the heart is at rest or between beats  
- Cholesterol: measured cholesterol  

The data set contains Weight, Diastolic Blood pressure, Systolic blood pressure and Cholesterol for alive subjects in the heart.csv.   


```{r}

# Read the CSV file
d1 <- read.csv("heart.csv") 

# Check data format
#str(d1)

```



The medical director wants to know if blood pressures and weight can better predict cholesterol outcome. Consider modeling cholesterol as a function of diastolic, systolic, and weight.   

***

#### Question (a): 
- Fit a linear regression model for cholesterol as a function of diastolic, systolic, and weight. Generate the diagnostics plots and comment on any issues that need to be noted. For Cook’s distances, do not leave any points that have Cook’s distance greater than 0.015.  
 
#### Answer -:  


##### Step 1) Run linear regressions with multicollinearity issue  
```{r}
e2.lr <- lm( Cholesterol ~ Diastolic+Systolic+Weight  , data=d1)
summary(e2.lr)

```
##### Result:   
- P-Value of F-Test is significant. Thus, the model is useful. 
        
In the other word, there exist linear regression for cholesterol as a function of diastolic, systolic, and weight.


##### Step 2) Perform Correlation to find Multicollinearity   
- Two methods:
        
          - 1) Pairwise Scatter Plot  
        
          - 2) Quantitative Method (VIF's)  

```{r}
pairs(d1,pch=19, col="darkblue")

```

##### Result:   
- It seems, maybe there exist a small correlation between Diastolic and Systolic, to be sure
we will perform a Quantitative test (VIF's)     

```{r}
vif(e2.lr)
```
##### Result:   
- There is no VIF>10, thus there exist no correlation between predictors.  


##### Step 3) Perform Diagnostic Plot    
```{r}
par(mfrow=c(2,2))
plot(e2.lr,which=1:4,col="darkblue")
```

##### Result:   
- It seems a normal distribution in Normal QQ plot, in Standardized residual, 95% data are between 0-1.5, thus data follow normal distribution.  
        
- In the Residuals plot, there is no pattern, thus there exist a Homoscedasticity.  
        
- In the cook distance, there exist some pints greater than 0.015  
        



##### Step 4) Influential points
```{r}
cook.d2 <- cooks.distance(e2.lr)
plot(cook.d2,col="darkblue",pch=19,cex=1)

```


- Delete observations larger than criteria (0.015)  

```{r}
e2.inf.id <- which(cooks.distance(e2.lr)>0.015)
e2.lr2 <- lm(Cholesterol ~ Diastolic+Systolic+Weight  , data=d1[-e2.inf.id,])
summary(e2.lr2)

```

##### Linear regression model:   

        ŷ = 156.32618 + 0.24922 * Diastolic + 0.30073 * Systolic + 0.03671 * Weight    

##### R Square (R2) and Adj. R Square: 

Calculate R2 and Adj.R2 
```{r}
summary(e2.lr2)$r.squared
summary(e2.lr2)$adj.r.squared
```
        R Square is very small, thus, "Goodness of fit" or "Predictive power" is very low.  
        
        Adj. R Square is very small too.  


##### Step 5) Plot scatter, with and without influential points   
```{r}
with(d1,plot(Cholesterol ~ Diastolic+Systolic+Weight, col="darkblue"))
abline(e2.lr,col="red")
abline(e2.lr2,col="green")
legend("bottomright",col=c("red","green"),legend = c("W/Inf. points","W/out Inf. points"),cex=0.8,title.adj = 0.15,lty=1)


```


##### Result:  
- Since dataset is very big (3134 observations) and we only remove 2 outliers, 
thus, the linear regression is closed together. The red line is under the green line.   


##### Step 6) Diagnostic plot for without influential points  

```{r}
par(mfrow=c(2,2))
plot(e2.lr2, which=c(1:4),col="darkblue") 
```


#### Conclusion:
Regression lines with/without influential points are almost the same.

***

#### Question (b): 
- Comment on the significance of the parameters and how much variation in cholesterol is described by the model. Comment on the relationship between cholesterol and statistically significant predictor(s). Check multicollinearity issue among predictors. Explain to the medical director whether this is a good model for the prediction of Cholesterol levels.   
 
#### Answer -:  


```{r}
summary(e2.lr2)
summary(e2.lr2)$r.squared
summary(e2.lr2)$adj.r.squared
```

##### Result:  
        - Diastolic and Systolic have significant value (Pvalue < 0.05)  
        - Weight has not significant value (Pvalue > 0.05) 

        - Based on this results, linear regression model:   

        ŷ = 156.32618 + 0.24922 * Diastolic + 0.30073 * Systolic + 0.03671 * Weight   

##### Check the multicollinearity:  
        in <step 2-part (a)>  we checked the multidisciplinary with these two methods:  
        - 1) Pairwise Scatter Plot  
        - 2) Quantitative Method (VIF's)  
        There is no VIF>10, thus there exist no correlation between predictors.  

#### Conclusion:   
        - Since P-Value is significant, there exist a linear regression for cholesterol as function of diastolic and systolic and weight
         although Pvalue of Weight is not significant, thus we must use a model selection to detect best model (Will do it in Exercise 3)  

        - R square and Adj. R Square are small, thus "Power of Prediction" or "Goodness of fit" is not good enough, 
        it means model is useful but is not good enough for prediction.  


***
